---
title: "Introduction to quickchartR"
output: rmarkdown::html_vignette
author: "Jacek Myna, Aleksandra Łuczak, Agata Pałdyna, Tomasz Radzikowski, Jan Sawicki"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{quickchartR-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(quickchartR)
```
#Introduction

Quickchart.io is a powerful application which provides fine and fast charts of different types. The problem is that the way of creating  charts is based on a specific URL form, what is not convenient for example big datasets. To improve functionality we have developed quickchartR (pronounce quickcharter), which enables using R language to create Quickchart objects.
In quickchartR You can create every type of chart from Quickchart.io except Sparklines and Mixed Charts ( [See why quickchartR does not support Sparklines](#why-quickchartr-does-not-support-sparklines)):

  * Bar Graph
  * Line Graph
  * Radar Chart
  * Pie Chart
  * Doughnut Chart
  * Scatter Plot
  * Bubbel Chart


&nbsp;

&nbsp;


# How to use quickchartR?


## presumptions

In the following examples we are working with widely known data from Iris dataset which is built in R. 




#FAQ

### why quickchartR does not support sparklines

prepared text
